Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2021-09-29 psy_peds_3.Rmd
graphic outputs added to the new version of the analysis

listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra")
paket(listOfPackages)
```

@@

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
```

@@

```{r}
icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
   select( disorder_group, ICD10_Code, description )
```

@@

```{r}
# Adjusting patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
last_data_date <- max(as.Date(demo_raw$last_discharge_date, "%Y-%m-%d"))
clinical_raw <- left_join(clinical_raw, demo_raw[c("patient_num", "cohort", "age", "sex")], by = c("patient_num", "cohort"))
clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)
clinical_raw <- select(clinical_raw, -age)

@@

# Steps necessary to filter:
  # - encounters without ICD
  # - ICD diagnosed outside of encounters
  # - patients from the patients table that don't have any encounters matching the criteria
clinical_table <- clinical_filtered %>% 
  inner_join( distinct(obs_filtered, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation) %>% 
  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
observation_table <- obs_filtered %>% 
  inner_join(distinct(clinical_table, patient_num, days_since_admission), by = c("patient_num", "days_since_admission"))
demographic_table <- demo_raw %>% 
  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
all_codes <- observation_table %>% 
  inner_join(clinical_table, on = c("patient_num", "days_since_admission")) %>% 
  inner_join(select(demographic_table, -days_since_admission), on = c("patient_num"))
stopifnot(nrow(all_codes) == nrow(observation_table))
stopifnot(unique(all_codes$cohort) == "AllAdm")
stopifnot(unique(all_codes$in_hospital) == 1)
stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))

@@

th_day <- ggplot(total_hospitalizations_per_day, aes(x = date,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Total admissions (per day)")

th_week <- ggplot(unique_patients_hospitalized_per_week, aes(x = weeks,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Total admissions (per week)")

@@

@@

all_codes_desc <- all_codes %>% 
    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>%
  dplyr::select( patient_num, cohort, sex, age_time_visit, concept_type, concept_code, calendar_date,
                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
                 period, disorder_group, description, icd_code_category )

@@

if( raceAvailable == TRUE ){
  all_codes_psy <- all_codes_psy %>%
    left_join( race_raw, by = c("patient_num", "cohort") )
}
```

@@

count_icd_per_day$percentage_psy <- count_icd_per_day$count_psy / count_icd_per_day$count_icd

thpd_psyPer <- count_icd_per_day %>%
  ggplot(aes(x = date, y = percentage_psy, fill = period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by day)",
       title = "Percentage of patients with mental health related ICD codes (daily)") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyPer
```

```{r}
thpd_psyCount <- count_icd_per_day %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = date, y = value, fill = counts, color = counts)) +
  geom_point() +
  # geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Date (by day)",
       title = "Daily patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount
```

@@

>>>>

listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr")
paket(listOfPackages)
```

@@

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "../Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
```

@@

```{r}
icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
   dplyr::select( disorder_group, ICD10_Code, description )
```

@@

```{r}
# Adjusting patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
last_data_date <- max(as.Date(demo_raw$last_discharge_date, "%Y-%m-%d"))
clinical_raw <- left_join(clinical_raw, demo_raw[c("patient_num", "cohort", "age", "sex")], by = c("patient_num", "cohort"))
clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)
clinical_raw <- dplyr::select(clinical_raw, -age)

@@

# Steps necessary to filter rows:
  # - encounters without ICD
  # - ICD diagnosed outside of encounters
  # - patients from the patients table that don't have any encounters matching the criteria
clinical_table <- clinical_filtered %>% 
  inner_join( distinct(obs_filtered, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation) %>% 
  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
# Adding new columns
observation_table <- obs_filtered %>% 
  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
demographic_table <- demo_raw %>% 
  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
all_codes <- observation_table %>% 
   inner_join(clinical_table, by = c("patient_num", "days_since_admission")) 
#%>% 
#   inner_join(dplyr::select(demographic_table, -days_since_admission), by = c("patient_num"))

stopifnot(unique(all_codes$cohort) == "AllAdm")
stopifnot(unique(all_codes$in_hospital) == 1)
stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
stopifnot(nrow(all_codes) == nrow(observation_table))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(demographic_table$patient_num)))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(observation_table$patient_num)))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(clinical_table$patient_num)))
stopifnot(nrow(distinct(all_codes, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table, patient_num, n_hospitalisation)))
stopifnot(nrow(distinct(all_codes, patient_num, days_since_admission)) == nrow(distinct(observation_table, patient_num, days_since_admission)))
```

@@

th_day <- ggplot(total_hospitalizations_per_day, aes(x = date,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Currently hospitalized patients (per day)")

th_week <- ggplot(unique_patients_hospitalized_per_week, aes(x = weeks,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Currently hospitalized patients (per week)")

@@

#### Group by sex

#### Per week
```{r}
# counts per week and sex
patients_hospitalized_per_week_agg_sex <- clinical_filtered %>%
  filter(date < end_date_plots &
         date >= start_date_plots & 
           sex %in% c("female", "male")) %>%
    group_by( weeks, sex ) %>%
  summarise(count_sex = n_distinct(patient_num))

# the total counts per week was estimated before: unique_patients_hospitalized_per_week
patients_hospitalized_per_week_agg_sex_perc <- unique_patients_hospitalized_per_week %>%
  left_join( patients_hospitalized_per_week_agg_sex, by = c("weeks")) %>% 
  replace_na(list(count_sex = 0))

patients_hospitalized_per_week_agg_sex_perc$percentage <- 100*patients_hospitalized_per_week_agg_sex_perc$count_sex/patients_hospitalized_per_week_agg_sex_perc$count

patients_hospitalized_per_week_agg_sex %>%
  ggplot(aes(x = weeks, y = count_sex, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Week",
       title = "Weekly patient counts aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)

patients_hospitalized_per_week_agg_sex_perc %>%
  ggplot(aes(x = weeks, y = percentage, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Percentage",
       x = "Week",
       title = "Weekly patient (%) aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
```

@@

all_codes_desc <- all_codes %>% 
    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>%
  dplyr::select( patient_num, sex, age_time_visit, concept_type, concept_code, calendar_date,
                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
                 period, disorder_group, description, icd_code_category )

@@

if( raceAvailable == TRUE ){
  all_codes_psy <- all_codes_psy %>%
    left_join( race_raw, by = c("patient_num") )
}
```

### Aggregated by sex
```{r}
# counts per week and sex
patients_hospitalized_per_week_agg_sex <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots & 
           sex %in% c("female", "male")) %>%
    group_by( weeks, sex ) %>%
  summarise(count_sex = n_distinct(patient_num))

patients_hospitalized_per_week_agg_sex %>%
  ggplot(aes(x = weeks, y = count_sex, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Week",
       title = "Weekly patient counts aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)

##percentage
total_patients_psy_per_week <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( weeks ) %>%
  summarise(count = n_distinct(patient_num))

patients_hospitalized_per_week_agg_sex_perc <- total_patients_psy_per_week %>%
  left_join( patients_hospitalized_per_week_agg_sex, by = c("weeks")) %>% 
  replace_na(list(count_sex = 0))

patients_hospitalized_per_week_agg_sex_perc$percentage <- 100*patients_hospitalized_per_week_agg_sex_perc$count_sex/patients_hospitalized_per_week_agg_sex_perc$count

patients_hospitalized_per_week_agg_sex_perc %>%
  ggplot(aes(x = weeks, y = percentage, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Percentage ",
       x = "Week",
       title = "Weekly patient (%) aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)

@@

count_icd_per_day$percentage_psy <- 100*count_icd_per_day$count_psy / count_icd_per_day$count_icd

thpd_psyPer <- count_icd_per_day %>%
  ggplot(aes(x = date, y = percentage_psy, fill = period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by day)",
       title = "Percentage of patients with mental health related ICD codes (daily)") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
  geom_point()
thpd_psyPer

count_icd_per_day %>%
  ggplot(aes(x = date, y = percentage_psy, fill = period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by day)",
       title = "Percentage of patients with mental health related ICD codes (daily)") + 
  geom_point()
  #geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
```

By week

```{r}
count_icd_per_week_psy <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( weeks ) %>%
  summarise(count_psy = n_distinct(patient_num)) %>%
  mutate( period = ifelse( weeks < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

count_icd_per_week <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( weeks ) %>%
  summarise(count_icd = n_distinct(patient_num)) %>%
  mutate( period = ifelse( weeks < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

count_icd_per_week <- count_icd_per_week %>%
  left_join(count_icd_per_week_psy, 
            by = c("weeks", "period")) %>% 
  replace_na(list(count_psy = 0))

count_icd_per_week$percentage_psy <- 100*count_icd_per_week$count_psy / count_icd_per_week$count_icd

thpd_psyPer <- count_icd_per_week %>%
  ggplot(aes(x = weeks, y = percentage_psy, fill=period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by week)",
       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
  geom_point()
thpd_psyPer

count_icd_per_week %>%
  ggplot(aes(x = weeks, y = percentage_psy, fill = period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by week)",
       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
  geom_point()
  #geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
```

# Per day
```{r}
thpd_psyCount <- count_icd_per_day %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = date, y = value, fill = counts, color = counts)) +
  geom_point() +
  # geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Date (by day)",
       title = "Daily patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount

# Per week
```{r}
thpd_psyCount <- count_icd_per_week %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = weeks, y = value, fill = counts, color = counts)) +
  geom_point() +
  # geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Date (by week)",
       title = "Weekly patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount

@@

#### Psychiatric conditions aggregating by suptype

##### Total hospitalizations

We estimate and plot the counts per disorder group.

Per day
```{r}
count_disorder_group_per_day <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( date, disorder_group ) %>%
  summarise(count_dg = n_distinct(patient_num)) %>%
  mutate( period = ifelse( date < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

dg_per_day <- count_disorder_group_per_day %>%  
  ggplot(aes(x = date, y= count_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Days",
       y = "Patient Count",
       title = "Daily patient counts with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dg_per_day
```

Per week
```{r}
count_disorder_group_per_week <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( weeks, disorder_group ) %>%
  summarise(count_dg = n_distinct(patient_num)) %>%
  mutate( period = ifelse( weeks < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

dg_per_week <- count_disorder_group_per_week %>%  
  ggplot(aes(x = weeks, y= count_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Weeks",
       y = "Patient Count",
       title = "Weekly patient counts with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dg_per_week
```

Per month
```{r}
count_disorder_group_per_month <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( month, disorder_group ) %>%
  summarise(count_dg = n_distinct(patient_num)) %>%
  mutate( period = ifelse( month < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

dg_per_month <- count_disorder_group_per_month %>%  
  ggplot(aes(x = month, y= count_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Months",
       y = "Patient Count",
       title = "Monthly patient counts with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dg_per_month
```

Percentage using as denominator total number of patients per day with a psy disorder.

Per day
- Total counts previously estimated (count_icd_per_day_psy).
- Counts per disorder group previously estimated (count_disorder_group_per_day)

```{r}
# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_day <- count_icd_per_day_psy %>%
  left_join(count_disorder_group_per_day, 
            by = c("date", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_day$percentage_dg <- 100*perc_disorder_group_per_day$count_dg / perc_disorder_group_per_day$count_psy

dgPerc_per_day <- perc_disorder_group_per_day %>%  
  ggplot(aes(x = date, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Days",
       y = "Patient (%)",
       title = "Percentage of daily patients with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_per_day
```

Per week
- Counts per disorder group previously estimated (count_disorder_group_per_week)

```{r}
# estimate the weekly counts
count_icd_per_week_psy <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( weeks ) %>%
  summarise(count_psy = n_distinct(patient_num)) %>%
  mutate( period = ifelse( weeks < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_week <- count_icd_per_week_psy %>%
  left_join(count_disorder_group_per_week, 
            by = c("weeks", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_week$percentage_dg <- 100*perc_disorder_group_per_week$count_dg / perc_disorder_group_per_week$count_psy

dgPerc_per_week <- perc_disorder_group_per_week %>%  
  ggplot(aes(x = weeks, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Weeks",
       y = "Patient (%)",
       title = "Percentage of weekly patients with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_per_week
```

Months 
```{r}
# estimate the monthly counts
count_icd_per_month_psy <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( month ) %>%
  summarise(count_psy = n_distinct(patient_num)) %>%
  mutate( period = ifelse( month < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_month <- count_icd_per_month_psy %>%
  left_join(count_disorder_group_per_month, 
            by = c("month", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_month$percentage_dg <- 100*perc_disorder_group_per_month$count_dg / perc_disorder_group_per_month$count_psy

dgPerc_per_month <- perc_disorder_group_per_month %>%  
  ggplot(aes(x = month, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Months",
       y = "Patient (%)",
       title = "Percentage of monthly patients with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_per_month
```

Percentage using as denominator total number of patients

Per day
- Counts per disorder group previously estimated (count_disorder_group_per_day)

```{r}
# estimate the total number per day 
count_icd_per_day <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( date ) %>%
  summarise(count_icd = n_distinct(patient_num)) %>%
  mutate( period = ifelse( date < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_day <- count_icd_per_day %>%
  left_join(count_disorder_group_per_day, 
            by = c("date", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_day$percentage_dg <- 100*perc_disorder_group_per_day$count_dg / perc_disorder_group_per_day$count_icd

dgPerc_per_day <- perc_disorder_group_per_day %>%  
  ggplot(aes(x = date, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Days",
       y = "Patient (%)",
       title = "Percentage of daily patients with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_per_day
```

Per week
- Counts per disorder group previously estimated (count_disorder_group_per_week)

```{r}
# estimate the weekly counts
count_icd_per_week <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( weeks ) %>%
  summarise(count_icd = n_distinct(patient_num)) %>%
  mutate( period = ifelse( weeks < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_week <- count_icd_per_week %>%
  left_join(count_disorder_group_per_week, 
            by = c("weeks", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_week$percentage_dg <- 100*perc_disorder_group_per_week$count_dg / perc_disorder_group_per_week$count_icd

dgPerc_per_week <- perc_disorder_group_per_week %>%  
  ggplot(aes(x = weeks, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Weeks",
       y = "Patient (%)",
       title = "Percentage of weekly patients with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_per_week
```

Months 
```{r}
# estimate the monthly counts
count_icd_per_month <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( month ) %>%
  summarise(count_icd = n_distinct(patient_num)) %>%
  mutate( period = ifelse( month < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_month <- count_icd_per_month %>%
  left_join(count_disorder_group_per_month, 
            by = c("month", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_month$percentage_dg <- 100*perc_disorder_group_per_month$count_dg / perc_disorder_group_per_month$count_icd

dgPerc_per_month <- perc_disorder_group_per_month %>%  
  ggplot(aes(x = month, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Months",
       y = "Patient (%)",
       title = "Percentage of monthly patients with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_per_month
```
